---
title: "2020-07-03 World university ranking"
author: "elvyna"
date: "03/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
DIRECTORY <- 'jupyter-notebook/dataset/world-university-ranking/'
knitr::opts_knit$set(root.dir = DIRECTORY)
```

## Data Preparation

Tidy up the data format.

```{r}
df.times <- read.csv('datasets_27_792993_timesData.csv')
country.mapping <- read.csv('country.csv')
df.times$international <- as.numeric(df.times$international)

library(stringr)
df.times$num_students <- as.numeric(str_replace(df.times$num_students,',',''))
df.times$total_score <- as.numeric(str_replace(df.times$total_score,'-',''))
df.times$world_rank <- as.numeric(str_replace(df.times$world_rank,'=',''))
df.times$international_students <- as.numeric(str_replace(df.times$international_students,'%','')) / 100
df.times$female.ratio <- as.numeric(sapply(str_split(df.times$female_male_ratio," : "), getElement, 1))

df.times$country <- as.character(df.times$country)
df.times$country <- ifelse(df.times$country == "Unisted States of America", "United States of America", 
                           ifelse(df.times$country == "Unted Kingdom", "United Kingdom", df.times$country)
                           )

df.times <- df.times %>% merge(country.mapping, how='left', by = c('country'='country'))
str(df.times)
head(df.times)

sort(unique(df.times$country))

table(df.times$year)

library(dplyr)
df.2016 <- df.times %>% filter(year == 2016)
```

## Observe normality

The numeric variables have different value ranges, and they are not normally distributed.

```{r}
library(ggplot2)
library(ggthemes)
theme_set(theme_clean())
library(reshape2)

numeric.vars <- c('teaching', 'international', 'research', 'citations',
                  'num_students', 'student_staff_ratio', 'international_students', 'female.ratio')
df.2016.melt <- df.2016 %>% melt(id.vars = c('world_rank', 'university_name','country'),
                 measure.vars = numeric.vars)

ggplot(data = df.2016.melt) +
  geom_histogram(aes(x = value)) +
  facet_wrap(.~variable, scales = 'free')

summary(df.2016)
```


## Principal Component Analysis

Let's reduce the dimensionality by retaining the linear combinations of variables that preserve most variability of the data. 

```{r}
## remove NA
df.2016.complete <- df.2016[complete.cases(df.2016[, numeric.vars]),]
pc <- prcomp(df.2016.complete[, numeric.vars], 
       scale = TRUE)
screeplot(pc)

print(paste("Variation represented in 2 components: ", sum((pc$sdev^2)[1:2]) / sum(pc$sdev^2)))
print(paste("Variation represented in 3 components: ", sum((pc$sdev^2)[1:3]) / sum(pc$sdev^2)))
```

There are some outliers: 
* Asia: Anadolu University
* Africa: University of South Africa, Cairo University

```{r}
library(scatterplot3d)
library(car)

scatterplot3d(
  x = pc$x[,1],
  y = pc$x[,2],
  z = pc$x[,3],
  color = as.numeric(as.factor(df.2016.complete$country))
)

library(plotly)
ggplot(data = df.2016.complete) +
  geom_point(aes(
    x = pc$x[,1],
    y = pc$x[,2],
    col = as.factor(continent),
    text = university_name
  ),
  alpha = 0.8) +
  scale_color_brewer(
    name = 'Continent',
    palette = 'Set2'
  ) + scale_x_continuous(
    name = 'PC1'
  ) + scale_y_continuous(
    name = 'PC2'
  ) + theme(
    legend.position = 'top',
    legend.direction = 'horizontal'
  )
ggplotly(tooltip = 'text')
```

## Rotated PCA

Aim: to get interpretable meaning of the loadings.

The first component is highly correlated with teaching, research, and citations. While the second component is highly correlated to number of students -- including the student-staff ratio and ratio of female vs. male. The third component is highly correlated to international outlook and number of international students.

Taking the rotated loadings into consideration, we know that the three outliers have quite large number of students and student-staff ratio; but their teaching, research, and citation values are lower than the average universities' (notice that they have negative PC1 values).
```{r}
pc.rotated <- varimax(pc$rotation[,1:3])
print(round(pc.rotated$loadings,2), cutoff=.2)
```

```{r}
pc1.loadings <- pc.rotated$loadings[,1]
pc2.loadings <- pc.rotated$loadings[,2]
pc.rotated.loadings <- cbind(as.data.frame(pc1.loadings), as.data.frame(pc2.loadings))
pc.rotated.loadings$original.variable <- row.names(pc.rotated.loadings)

pc.rotated.loadings.melt <- pc.rotated.loadings %>% melt(
  id.vars = 'original.variable',
  measure.vars = c('pc1.loadings','pc2.loadings')
)

ggplot(data = pc.rotated.loadings.melt) + geom_bar(
  aes(
    x = original.variable,
    y = value,
    fill = variable
    ), 
  stat = 'identity',
  position = 'dodge'
  ) + scale_fill_brewer(
    palette = 'Set2'
  ) + scale_x_discrete(
    name = ''
  ) + scale_y_continuous(
    name = 'correlation',
    breaks = seq(-1,1,.1)
  ) + geom_hline(
    yintercept = .2,
    color = 'blue',
    linetype = 'dashed'
  ) + geom_text(
    aes(
      x = 8.5,
      y = .3,
      label = 'threshold: 0.2'
      ), 
    colour = 'blue'
    ) + coord_flip()
```
